---
title: "Penguin Size Comparisons"
author: "Anna Talken"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms

ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## Quantile-quantile plot to check normality

ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)


## Get descriptive statistics
## Creating a summary table but not storing it

chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

```


Get vectors of body_mass_g for female and male chinstrap penguins separately
Using pull to get vector values

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest #Calling the t-test to view output

cohen.d(chinstraps_f, chinstraps_m) # looks for effect size, in this case there is a LARGE effect size - pretty clear difference between female and male chinstrap body size
```

Are male chinstrap penuins significantly larger (body mass) than female chinstrap penguins?

```{r}
# Need to do a one-sided t-test
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # order of vectors and values matters: are chinstraps body mass for MALE greater than FEMALE
```

## One-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguins_counts <- penguins %>% 
  count(species, sex) #Making a table of number of penguins by species and sex
penguins_counts
```

Create a subset of only female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) +
  facet_wrap(~species, scales = "free") # gives each graph its own scale on the y-axis

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")
```

### Summary statistics

Starting from penguins_f, create a summary table that groups by species and returns the following:
- mean of the flipper length
- sd
- sample size
- variance 'var'
- standard error = sd/sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper = mean(flipper_length_mm),
    sd_flipper = sd(flipper_length_mm),
    variance_flipper = var(flipper_length_mm),
    sample_size = n(),
    se_flipper = sd(flipper_length_mm)/sqrt(n())
  )

# Levene test for equal variance.... but even if there is a significant difference with a Levene test, it could just be because the sample size is so large

leveneTest(flipper_length_mm ~ species, data = penguin_f)

# 0.05782 is greater than the 0.05 level, so we retain the NULL hypothesis
```

### Beeswarm plot with mean +- 1sd

Shows distribution of observations

```{r}
ggplot() + 
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.9) + # transparency
geom_point(data = penguin_f_summary, aes(x = species, y = mean_flipper),
           size = 4) +
  geom_errorbar(data = penguin_f_summary, 
                aes( x = species, 
                     ymax = mean_flipper + sd_flipper,
                     ymin = mean_flipper - sd_flipper),
                width = 0.1) + # chnages width of the error bar
  theme_light()
```

### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov) # displays the anova below

# Large F-stat means that the variance between groups is very large compared to the variance within groups....reject null hypothesis

TukeyHSD(penguin_aov) #pairwise adjusted p-value for each two pair comparison
```

Type III ANOVA with the car package

Use Type III if you have an unbalanced design

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f) # Create a linear model
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```











