---
title: "Penguin Size Comparisons"
author: "Anna Talken"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms

ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## Quantile-quantile plot to check normality

ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)


## Get descriptive statistics
## Creating a summary table but not storing it

chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

```


Get vectors of body_mass_g for female and male chinstrap penguins separately
Using pull to get vector values

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest #Calling the t-test to view output

cohen.d(chinstraps_f, chinstraps_m) # looks for effect size, in this case there is a LARGE effect size - pretty clear difference between female and male chinstrap body size
```

Are male chinstrap penuins significantly larger (body mass) than female chinstrap penguins?

```{r}
# Need to do a one-sided t-test
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # order of vectors and values matters: are chinstraps body mass for MALE greater than FEMALE
```

















